Global study of well-being and internet adoption | Online analysis supplement

### 2021-11-15 01-clean.Rmd
clean analyses

```{r packages}
#| results: 'hide'
# Packages
library(readxl)
library(imputeTS)
library(naniar)
library(janitor)
library(countrycode)
library(kableExtra)
library(tidyverse)

@@

## Data cleaning

We downloaded the relevant data from sources to `data-raw/`. We cannot share the data here but indicate below how and when each data table was obtained.

### Global Burden of Disease

@@

### Gallup World Poll

@@

### International Telecommunication Union

The internet adoption data were downloaded from the ITU website in October 2021, and placed with their original file names to `data-raw/ITU/`. The specific URLs for obtaining the data files are

- <https://www.itu.int/en/ITU-D/Statistics/Documents/statistics/2021/July/PercentIndividualsUsingInternet.xlsx>
- <https://www.itu.int/en/ITU-D/Statistics/Documents/statistics/2021/July/FixedBroadbandSubscriptions_2000-2020.xlsx>
- <https://www.itu.int/en/ITU-D/Statistics/Documents/statistics/2021/July/MobileBroadbandSubscriptions_2007-2020.xlsx>

We first load, reshape, and merge these datasets.

```{r itu-load-clean}
# Read and reshape fixed broadband data
itu_fixed <- read_xlsx(
  "data-raw/ITU/FixedBroadbandSubscriptions_2000-2020.xlsx", 
  sheet = 2,
  col_types = "text"
) %>% 
  pivot_longer(-c(Indicator, Country)) %>% 
  separate(name, into = c("year", "variable")) %>% 
  pivot_wider(names_from = variable) %>% 
  mutate(fixed = as.numeric(value)) %>% 
  select(country = Country, year, fixed)

@@

#### Missingness

@@

There are some missing values in the data; in the Internet variable these are mainly in the first or last years of each country's time-series. This needs to be changed to calculate by-variable (with appropriate time-spans) when adding mobile and fixed

```{r}
miss_var_summary(itu)
```

We then impute missing values to the internet adoption timeseries. We use linear interpolation to fill missing intermediate values, but do not extrapolate before or after the first and last values.

```{r itu-impute}
itu <- itu %>% 
  group_by(country) %>% 
  # Find first and last year with internet data in each country
  mutate(
    min_year = year[min(which(!is.na(internet)))],
    max_year = year[max(which(!is.na(internet)))]
  ) %>%
  # Interpolate values linearly
  # use possibly() to avoid errors for countries with no data
  # Add variables to c() to impute for fixed and mobile broadband
  mutate(
    across(
      c(internet), 
      possibly(
        function(x) na_interpolation(x, maxgap = Inf), 
        # If function errors, provide original internet value
        otherwise = internet
      )
    )
  ) %>% 
  # Take out projected (non-intermediate) imputed values
  mutate(
    internet = if_else(
      between(year, unique(min_year), unique(max_year)), 
      internet, 
      NaN
    )
  ) %>% 
  ungroup() %>% 
  select(-min_year, -max_year)
```

We see that 96 intermediate internet use values were imputed to the time series.

```{r}
miss_var_summary(itu)
```

### Complete study dataset

#### GLOBE regions

@@

# Load our manually filled region data
regions_filled <- read_csv(
  "data-raw/Regions/GBD-countries-without-region-filled.csv"
) 
```

@@

#### Create predictors

We then create predictors (centered, lagged, rescaled) for models in the ITU dataset.

```{r}
# For models, we rescale year; centred at 2013 and one unit indicates decade
itu <- itu %>% 
  mutate(time = (year - 2013) / 10)

# Create predictor: Lagged effect of 10% increase in internet
itu <- itu %>% 
  arrange(country, year) %>% 
  group_by(country) %>% 
  mutate(i1 = lag(internet*10, 1)) %>% 
  ungroup()

# Centering within and between countries
itu <- itu %>%
  bmlm::isolate("i1", by = "country", which = "both")
```

#### Create tables

@@

We also save a synthetic version of the gwp dataset to better enable others to reproduce our analytical procedures:

```{r synth-data-create}

```

>>>>

```{r packages}
#| results: 'hide'
# Packages
library(readxl)
library(imputeTS)
library(naniar)
library(scales)
library(janitor)
library(countrycode)
library(kableExtra)
library(tidyverse)

@@

## Global Burden of Disease

@@

## Gallup World Poll

@@

## International Telecommunication Union

We first download the data files from the ITU website (in November 2021). Note that new versions may be posted subsequent to this, in which case the URLs would change and should then be changed here too to update the analyses with latest data.

```{r itu-download}
# Download data sets from the ITU website if not yet downloaded
if (!file.exists("data-raw/ITU/PercentIndividualsUsingInternet.xlsx")) {
  dir.create("data-raw/ITU")
  download.file(
    "https://www.itu.int/en/ITU-D/Statistics/Documents/statistics/2021/PercentIndividualsUsingInternet_Nov2021.xlsx", 
    "data-raw/ITU/PercentIndividualsUsingInternet.xlsx"
  )
  download.file(
    "https://www.itu.int/en/ITU-D/Statistics/Documents/statistics/2021/July/FixedBroadbandSubscriptions_2000-2020.xlsx",
    "data-raw/ITU/FixedBroadbandSubscriptions_2000-2020.xlsx"
  )
  download.file(
    "https://www.itu.int/en/ITU-D/Statistics/Documents/statistics/2021/July/MobileBroadbandSubscriptions_2007-2020.xlsx",
    "data-raw/ITU/MobileBroadbandSubscriptions_2007-2020.xlsx"
  )
}
```

We then read and reshape the ITU data

```{r itu-clean}
# Read and reshape fixed broadband data
itu_fixed <- read_xlsx(
  "data-raw/ITU/FixedBroadbandSubscriptions_2000-2020.xlsx", 
  sheet = 2,
  col_types = "text"
) %>% 
  pivot_longer(-c(Indicator, Country)) %>% 
  separate(name, into = c("year", "variable")) %>% 
  pivot_wider(names_from = variable) %>% 
  mutate(fixed = as.numeric(value)) %>% 
  select(country = Country, year, fixed)

@@

### Missingness

@@

We then impute missing values to the internet adoption timeseries. We use linear interpolation to fill missing intermediate values, but do not extrapolate before or after the first and last values of each country.

```{r itu-impute}
# Reshape to easily impute all variables
itu_long <- itu %>%
  pivot_longer(c(internet, fixed, mobile))

# Impute by country and variable
itu_long <- itu_long %>%
  group_by(country, name) %>% 
  # Find first and last year with this variable in each country
  # so as to prevent extrapolation
  mutate(
    min_year = year[min(which(!is.na(value)))],
    max_year = year[max(which(!is.na(value)))]
  ) %>%
  # Interpolate values linearly
  # use possibly() to avoid errors for countries with no data
  mutate(
    across(
      value, 
      .fns = list(
        i = possibly(
          function(x) na_interpolation(x, maxgap = Inf), 
          otherwise = NaN
        )
      )
    )
  ) %>% 
  # Take out projected (non-intermediate) imputed values
  mutate(
    value_i = if_else(
      between(year, unique(min_year), unique(max_year)), 
      value_i, 
      NaN
    )
  ) %>% 
  ungroup() %>% 
  select(-min_year, -max_year)
```

Here is a summary of how many values were imputed

```{r}
itu_long %>%   
  group_by(name) %>% 
  mutate(min_year = if_else(name == "mobile", 2007, 2000)) %>% 
  summarise(
    n_miss_orig = sum(is.na(value) & year >= min_year),
    p_miss_orig = percent(n_miss_orig / sum(year >= min_year), .1),
    n_miss_imp = sum(is.na(value_i) & year >= min_year),
    p_miss_imp = percent(n_miss_imp / sum(year >= min_year), .1),
    imputed = n_miss_orig - n_miss_imp
  )
```

We then replace the original values with the imputed ones

```{r}
itu <- itu_long %>% 
  mutate(value = value_i) %>% 
  select(-value_i) %>% 
  pivot_wider()
rm(itu_long)
```

## GLOBE regions

@@

# Load our manually filled region data
regions_filled <- read_csv(
  "data-raw/Regions/ITU-countries-without-region-filled.csv"
) 
```

@@

## Join tables

@@

### 2021-11-15 03-time.Rmd
clean analyses

@@

We first fit the models using lme4 because they are big and therefore would be very slow with brms.

@@

#### lme4

```{r fit-gbd-time-lme4}
# These take about 5-20 minutes each
# run in parallel in background jobs
# recommended to conduct this in interactive mode (rstudio)
file_path <- "output/lme4-gbd-time.rds"
if (!file.exists(file_path)) {
  library(job)
  job({fit_anx <- lmer(
    Anxiety ~ 1 + time * sex * age + 
      (1 + time * sex * age | region) +
      (1 + time * sex * age | country),
    data = gbd_wide
  )}, title = "Anxiety time (lmer)")
  job({fit_dep <- lmer(
    Depression ~ 1 + time * sex * age + 
      (1 + time * sex * age | region) +
      (1 + time * sex * age | country),
    data = gbd_wide
  )}, title = "Depression time (lmer)")
  job({fit_sel <- lmer(
    Selfharm ~ 1 + time * sex * age + 
      (1 + time * sex * age | region) +
      (1 + time * sex * age | country),
    data = gbd_wide
  )}, title = "Self harm time (lmer)")
  fits_lmer <- distinct(gbd, cause) %>% 
    mutate(fit = list(fit_anx, fit_dep, fit_sel))
  write_rds(fits_lmer, file_path)
} else {fits_lmer <- read_rds(file_path)}
```

```{r}
# Fixed effects parameters
fixed_effects <- fits_lmer %>% 
  mutate(out = map(fit, ~tidy(., conf.int = TRUE))) %>% 
  select(-fit) %>% 
  unnest(out) %>% 
  standardize_names() %>% 
  as_tibble() %>% 
  filter(
    Effects == "fixed", 
    Parameter != "(Intercept)"
  ) 
# Fixed effects per demographic
fixed_effects_demographics <- fits_lmer %>% 
  mutate(
    out = map(
      fit,
      ~emtrends(.x, var = "time", by = c("sex", "age")) %>% 
        parameters() %>% 
        as.data.frame()
    )
  ) %>% 
  select(-fit) %>% 
  unnest(out) %>% 
  standardize_names() %>% 
  as_tibble() %>% 
  mutate(Parameter = str_glue("{sex}\n({age})"))

@@

>>>>

```{r}
itu <- read_rds("data/itu.rds")
gbd <- read_rds("data/gbd.rds") # 10 to 24 year olds
```

## Reshape data for models

### Create predictors

We then create predictors (centered, lagged, rescaled) for models in the ITU dataset.

```{r}
# For models, year is centred at 
itu <- itu %>% 
  mutate(time = (year - 2013) / 10)

# Create predictor: Lagged effect of 10% increase in internet
itu <- itu %>% 
  arrange(country, year) %>% 
  group_by(country) %>% 
  mutate(i1 = lag(internet*10, 1)) %>% 
  ungroup()

# Centering within and between countries
itu %>%
  bmlm::isolate("i1", by = "country", which = "both")

tibble(
  id = c(1,1,1,2,2,2,3,3,3),
  x = c(1,2,3,4,5,6,7,8,9)
) %>% 
  bmlm::isolate("x", by = "id", which = "both") %>% 
  group_by(id) %>% 
  mutate(x_cb2 = mean(x), x_cw2 = x - x_cb2) %>% 
  ungroup() %>% 
  mutate(x_cb2 = x_cb2 - mean(x))
```

```{r}
# Contrast code sex
gbd$sex <- factor(gbd$sex)
contrasts(gbd$sex) <- c(-.5, .5)

# [todo] set mid age category as reference

# GBD in wide format
gbd_wide <- gbd %>%
  pivot_wider(names_from = cause, values_from = c(val, se))
names(gbd_wide) <- str_remove(names(gbd_wide), "val_")

# Transform edge proportions to fit beta distribution
gbd_10 <- gbd_wide %>% 
  mutate(internet = if_else(internet == 0, .0000000001, internet)) %>% 
  mutate(internet = if_else(internet == 1, .9999999999, internet))
```

@@

@@

@@

[todo] only pick the itu data for the internet model
